---
title: "Logistic Regression"
output: html_notebook
---

### **Research question : HR wants to identify the likelihood(how likely) of an employee leaves the organization**


```{r, echo=FALSE, warnings=FALSE}
library(openxlsx)
HR_Analytics <- read.xlsx("H:/Yashwanth/Kaggle/HR_comma_sep.xlsx",sheet = 1)
HR_Analytics$left <- factor(HR_Analytics$left, labels = c("Retain","left"))
HR_Analytics$Work_accident <- as.factor(HR_Analytics$Work_accident)
HR_Analytics$promotion_last_5years <- as.factor(HR_Analytics$promotion_last_5years)
```
---

#### **Proportion tables**
```{r, echo=FALSE, warnings=FALSE}
print(rbind(left=prop.table(table(HR_Analytics$left))*100,                              
      Work_accident=prop.table(table(HR_Analytics$Work_accident))*100,
      promotion_last_5years=prop.table(table(HR_Analytics$promotion_last_5years))*100))
      
print(rbind(salary=prop.table(table(HR_Analytics$salary,HR_Analytics$left))*100))

print(rbind(Department=prop.table(table(HR_Analytics$Department,HR_Analytics$left))*100))
```
**Insights**

* ~24% of employees has left organization
* ~14% of employees had Work accident
* ~2% of employees got promoted
* ~ 14% of employees from Low salary & ~9% of employees from medium salary has left organization
* ~7%, 4% and 5% of employees from Sales, Support & Technical department has left organization respectively

---

#### **Distribution of Variables : Histogram**

```{r, echo=FALSE, warning=FALSE}
library(ggplot2)
attach(HR_Analytics)
ggplot(HR_Analytics,aes(x=satisfaction_level)) +
  facet_wrap(~ left)+
  geom_histogram(aes(y=..density..),col = "blue2",bins = 40) +
  stat_function(fun = dnorm,args = list(mean=mean(last_evaluation),sd=sd(last_evaluation)),colour = "red") +
  labs(title = "Distribution of Satisfaction level", x = "left", y = "Satisfaction level") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```

```{r, echo=FALSE, warnings=FALSE}
library(psych)
print(describeBy(HR_Analytics$satisfaction_level,group = HR_Analytics$left)[1:2])
```

* Distribution of Satisfaction levels across employees who retained & left in/from the organization appears normal

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics,aes(x=last_evaluation)) +
  facet_wrap(~ left)+
  geom_histogram(aes(y=..density..),col = "blue2", bins = 40) +
  stat_function(fun = dnorm,args = list(mean=mean(last_evaluation),sd=sd(last_evaluation)),colour = "red") +
  labs(title = "Distribution of Last Evaluation", x = "left", y = "Last Evaluation") +
  theme(plot.title = element_text(hjust = 0.50))
detach(HR_Analytics)
```

```{r, echo=FALSE, warnings=FALSE}
print(describeBy(HR_Analytics$last_evaluation,group = HR_Analytics$left)[1:2])
```

* Distribution of Last Evaluation across employees who retained & left in/from the organization appears normal

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics,aes(x=number_project)) +
  facet_wrap(~ left)+
  geom_histogram(aes(y=..density..),col = "blue2", bins = 20) +
  stat_function(fun = dnorm,args = list(mean=mean(number_project),sd=sd(number_project)),colour = "red") +
  labs(title = "Distribution of Number of Projects", x = "left", y = "Number of Project") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```

```{r, echo=FALSE, warnings=FALSE}
print(describeBy(HR_Analytics$number_project,group = HR_Analytics$left)[1:2])
```

* Distribution of Number of Projects across employees who retained & left in/from the organization appears normal

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics,aes(x=average_monthly_hours)) +
  facet_wrap(~ left)+
  geom_histogram(aes(y=..density..),col = "blue2", bins = 40) +
  stat_function(fun = dnorm,args = list(mean=mean(average_monthly_hours),sd=sd(average_monthly_hours)),
                colour = "red") +
  labs(title = "Distribution of Average Monthly Hours", x = "left", y = "Average Monthly Hours") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```

```{r, echo=FALSE, warnings=FALSE}
print(describeBy(HR_Analytics$average_monthly_hours,group = HR_Analytics$left)[1:2])
```

* Distribution of Average Monthly hours across employees who retained & left in/from the organization appears normal

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics,aes(x=time_spend_company)) +
  facet_wrap(~ left)+
  geom_histogram(aes(y=..density..),col = "blue2", bins = 40) +
  stat_function(fun = dnorm,args = list(mean=mean(time_spend_company),sd=sd(time_spend_company)),colour = "red") +
  labs(title = "Distribution of Time Spend in Company", x = "left", y = "Time Spend in Company") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```

```{r, echo=FALSE, warnings=FALSE}
print(describeBy(HR_Analytics$time_spend_company,group = HR_Analytics$left)[1:2])
```

* Distribution of Time Spend in Company of the employees who retained appears right skewed & who left appears normal

---

#### **Distribution of Variables : Boxplot**

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics,aes(left,satisfaction_level)) +
  geom_boxplot() +
  labs(title = "Distribution of Satisfaction level", x = "left", y = "Satisfaction level") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```

* Distribution of Satisfaction levels across employees who retained & left in/from the organization appears normal

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics,aes(left,last_evaluation)) +
  geom_boxplot() +
  labs(title = "Distribution of Last Evaluation", x = "left", y = "Last Evaluation") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```
* Distribution of Last Evaluation across employees who retained & left in/from the organization appears normal

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics,aes(left,number_project)) +
  geom_boxplot() +
  labs(title = "Distribution of Number of Projects", x = "left", y = "Number of Project") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```
* 50th & 75th percentile of employees who retained shares same number of projects with one outlier(6)
* Distribution of Number of Projects across employees who left from the organization appears normal

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics,aes(left,average_monthly_hours)) +
  geom_boxplot() +
  labs(title = "Distribution of Average Monthly Hours", x = "left", y = "Average Monthly Hours") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```
* Distribution of Average Monthly Hours across employees who retained & left in/from the organization appears normal

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics,aes(left,time_spend_company)) +
  geom_boxplot() +
  labs(title = "Distribution of Time Spend in Company", x = "left", y = "Time Spend in Company") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```
* Distribution of Average Monthly Hours across employees who retained appeared right skewed with outliers
* Distribution of Average Monthly Hours across employees who left appears normal

---

#### **Relationship between Variables : Scatterplot**

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics, aes(last_evaluation,satisfaction_level)) +
  facet_wrap(~left) +
  geom_point(aes(color=left)) +
  geom_abline(color = "red") +
  labs(title="Relationship between Satisfaction level & Last Evaluation", 
       x = "last evaluation", y = "satisfaction level") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```
```{r, echo=FALSE, warnings=FALSE}
library(data.table)
HR_Analytics <- data.table(HR_Analytics)
print(HR_Analytics[,.(cor(satisfaction_level, last_evaluation)), by = left])
HR_Analytics <- data.frame(HR_Analytics)
print(cor(HR_Analytics$satisfaction_level,HR_Analytics$last_evaluation))
```

* There is Weak relationship between Satisfaction level & Last evaluation across employees who retained & left in/from the Organization

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics, aes(scale(HR_Analytics$number_project)[,1],scale(HR_Analytics$satisfaction_level)[,1])) +
  facet_wrap(~left) +
  geom_point(aes(color=left)) +
  geom_abline(color = "red") +
  labs(title="Relationship between Satisfaction level & Number of Projects", 
       x = "Number of Projects", y = "satisfaction level") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```

```{r, echo=FALSE, warnings=FALSE}
HR_Analytics <- data.table(HR_Analytics)
print(HR_Analytics[,.(cor(satisfaction_level, number_project)), by = left])
HR_Analytics <- data.frame(HR_Analytics)
print(cor(HR_Analytics$satisfaction_level,HR_Analytics$number_project))
```

* There is Weak relationship between Satisfaction level & Number of Projects across employees who retained & left in/from the Organization

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics, aes(scale(average_monthly_hours)[,1],scale(satisfaction_level)[,1])) +
  facet_wrap(~left) +
  geom_point(aes(color=left)) +
  geom_abline(color = "red") +
  labs(title="Relationship between Satisfaction level & Average Monthly Hours", 
       x = "average monthly hours", y = "satisfaction level") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```

```{r, echo=FALSE, warnings=FALSE}
HR_Analytics <- data.table(HR_Analytics)
print(HR_Analytics[,.(cor(satisfaction_level, average_monthly_hours)), by = left])
HR_Analytics <- data.frame(HR_Analytics)
print(cor(HR_Analytics$satisfaction_level,HR_Analytics$average_monthly_hours))
```

* There is Weak relationship between Satisfaction level & Average Monthly Hours across employees who retained & left in/from the Organization

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics, aes(scale(time_spend_company)[,1],scale(satisfaction_level)[,1])) +
  facet_wrap(~left) +
  geom_point(aes(color=left)) +
  geom_abline(color = "red") +
  labs(title="Relationship between Satisfaction level & Time Spend in Company", 
       x = "time spend in company", y = "satisfaction level") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```

```{r, echo=FALSE, warnings=FALSE}
HR_Analytics <- data.table(HR_Analytics)
print(HR_Analytics[,.(cor(satisfaction_level, time_spend_company)), by = left])
HR_Analytics <- data.frame(HR_Analytics)
print(cor(HR_Analytics$satisfaction_level,HR_Analytics$time_spend_company))
```

* There is Weak relationship between Satisfaction level & Time Spend across employees who retained in the Organization
* There is fairly strong relationship between Satisfaction level & Time Spend across employees who left from Organization

---

#### **Relationship between Variables(Correlation Analysis) : Tetrachoric & Polychoric Correlation** [Reference](http://john-uebersax.com/stat/tetra.htm#tsoft)

```{r, echo=FALSE, warnings=FALSE}
library(polycor)
polychor(table(HR_Analytics$left,HR_Analytics$Work_accident), ML = T, std.err = T)
```
* There is fairly negative relationship between employees who left & work accident

---

```{r, echo=FALSE, warnings=FALSE}
polychor(table(HR_Analytics$left,HR_Analytics$promotion_last_5years), ML = T, std.err = T)
```
* There is fairly negative relationship between employees who left & given promotion in last 5-years

---

```{r, echo=FALSE, warnings=FALSE}
polychor(table(HR_Analytics$left,HR_Analytics$salary), ML = T, std.err = T)
```
* There is weak negative relationship between employees who left & salary break-up

---

#### **Correlation Matrix**

```{r, echo=FALSE, warnings=FALSE}
cor(HR_Analytics[sapply(HR_Analytics,is.numeric)])
```
* In general, there is no strong correlation exists between variables

---

#### **Logistic Regression**

##### **Simple/Stratified Random Sampling**

```{r, echo=FALSE, warnings=FALSE}
library(sampling)
HR_Analytics <- cbind(HR_Analytics,row_id=row.names(HR_Analytics))
HR_Analytics_FT <- data.frame(table(HR_Analytics$left))
HR_Analytics_FT$Per <- (HR_Analytics_FT$Freq/sum(HR_Analytics_FT$Freq))*100
names(HR_Analytics_FT)[1] <- "left"

# Consider 70% of the data as sample
HR_Analytics_FT$Strata_Size <- ceiling((HR_Analytics_FT$Freq*(ceiling((dim(HR_Analytics)[1]/100)*70)/sum(HR_Analytics_FT$Freq))))
HR_Analytics_FT <- with(HR_Analytics_FT,HR_Analytics_FT[order(Strata_Size,decreasing = TRUE),])

# Stratification
HR_Analytics_Strata <- strata(HR_Analytics,c("left"),size = c(2500,8001), method = "srswor")
HR_Analytics_StRS <- getdata(HR_Analytics,HR_Analytics_Strata)
head(HR_Analytics_StRS)

# Test data
library(sqldf)
HR_Analytics_StRS_Test <- sqldf("select a.* from HR_Analytics a 
                                left join HR_Analytics_StRS b on a.row_id=b.row_id
                                where b.row_id is NULL")
```

---

##### **Logit Model** 

```{r, echo=FALSE, warnings=FALSE}
# Logit Model
HR_Logit <- glm(left ~ satisfaction_level+last_evaluation+number_project+average_monthly_hours
                        +Work_accident+promotion_last_5years+salary,
                        data = HR_Analytics_StRS, family = binomial)

# Odds ration & CI
cbind(OR = exp(cbind(coef(HR_Logit)))[,1]) #, confint(HR_Logit))
exp(cbind(OR = coef(HR_Logit), confint(HR_Logit)))
```

##### **Insights**
* For one unit increase in satisfaction level, odds of retain will increases by 0.01435192/0.014%
* For one unit increase in last evaluation, odds of retain will increases by 2.46506217/146%
* For one unit increase in number of projects, odds of retain will increases by 0.77209996/0.772%
* For one unit increase in last evaluation, odds of retain will increases by 1.00464547/4%

* For work accident1, odds of retain will increases by 0.22691262/ 0.226%
* For promotion in last 5yrs, odds of retain will increases by 0.30721556/ 0.307%
* For low-salary bucket, odds of retain will increases by 6.32788947/ 532%
* For low-salary bucket, odds of retain will increases by 4.02270236/ 302%

---

##### **Logit Model Validation**

```{r, echo=FALSE, warning=FALSE}
predict(HR_Logit,newdata=HR_Analytics_StRS_Test,type="response")
HR_OR_Pred <- data.frame(HR_Analytics_StRS_Test,actual=HR_Analytics_StRS_Test$left,
            prob=round(predict(HR_Logit,newdata=HR_Analytics_StRS_Test,type="response"),2),row.names=NULL)
HR_OR_Pred$OddsRatio <- round(HR_OR_Pred$prob/(1-HR_OR_Pred$prob),2)
HR_OR_Pred$Odds <- round(ifelse(HR_OR_Pred$OddsRatio<=1,HR_OR_Pred$OddsRatio*100,(HR_OR_Pred$OddsRatio*100)-100),2)
#predict(My_LHR_Logitogit,newdata=HR_Analytics_StRS_Test,type="terms")
HR_OR_Pred$Predict <- as.factor(ifelse(HR_OR_Pred$prob>=0.5,"left","Retain"))

# COnfusion Matrix
table(HR_OR_Pred$left)
table(Actual=HR_OR_Pred$left,Pred=HR_OR_Pred$Predict)

# Validation
library(caret)
HR_Accuracy <- (table(Actual=HR_OR_Pred$left,Pred=HR_OR_Pred$Predict)[3]+
                  table(Actual=HR_OR_Pred$left,Pred=HR_OR_Pred$Predict)[2])/nrow(HR_OR_Pred)
HR_Recall <- sensitivity(HR_OR_Pred$Predict,HR_OR_Pred$left)
HR_Precision <- posPredValue(HR_OR_Pred$Predict,HR_OR_Pred$left) 
```

##### **Insigths**
* Accuracy
* Recall
* Precision

---

#### **ROC-curve**
```{r, echo=FALSE, warning=FALSE}
dev.off()
library(pROC)
ROC <- roc(HR_OR_Pred$left,HR_OR_Pred$prob,plot=TRUE)                     # True versus True
# matrix(c(TP=ROC$sensitivities,TN=ROC$specificities,
#          FP=1-ROC$sensitivities,FN=1-ROC$specificities),nrow=3,ncol=4)

library(ROCR)
Prediction <- prediction(HR_OR_Pred$prob,HR_OR_Pred$left)
Perf1 <- performance(Prediction,measure="tpr",x.measure="fpr")    # True versus False
plot(Perf1,main="ROC Curve")

Perf2 <- performance(Prediction,measure="sens",x.measure="spec") # True versus True
plot(Perf2,main="Sensitivity & Specificity")
```


