---
title: "Linear Regression"
output: html_notebook
---

### **Research question : HR wants to evaluate how satisfied employees working for their Organization in order to define better employee retention policies**

#### **Hypothesis: Is there any significant difference between dependent & independent variables**
##### 	Null Hypothesis: There is NO significant difference between Satisfaction level & number of projects, average monthly hours, time spend in company, work accident, promotion last year, last evaluation
##### 	Alternate Hypothesis: There is a significant difference between Satisfaction level & number of projects, average monthly hours, time spent in company, work accident, promotion last year, last evaluation

```{r, echo=FALSE, warnings=FALSE}
library(openxlsx)
HR_Analytics <- read.xlsx("H:/Yashwanth/Kaggle/HR_comma_sep.xlsx",sheet = 1)
HR_Analytics$left <- factor(HR_Analytics$left, labels = c("Retain","Left"))
HR_Analytics$Work_accident <- as.factor(HR_Analytics$Work_accident)
HR_Analytics$promotion_last_5years <- as.factor(HR_Analytics$promotion_last_5years)
```
---

#### **Proportion tables**
```{r, echo=FALSE, warnings=FALSE}
print(rbind(left=prop.table(table(HR_Analytics$left))*100,                              
      Work_accident=prop.table(table(HR_Analytics$Work_accident))*100,
      promotion_last_5years=prop.table(table(HR_Analytics$promotion_last_5years))*100))
      
print(rbind(salary=prop.table(table(HR_Analytics$salary,HR_Analytics$left))*100))

print(rbind(Department=prop.table(table(HR_Analytics$Department,HR_Analytics$left))*100))
```
**Insights**

* ~24% of employees has left organization
* ~14% of employees had Work accident
* ~2% of employees got promoted
* ~ 14% of employees from Low salary & ~9% of employees from medium salary has left organization
* ~7%, 4% and 5% of employees from Sales, Support & Technical department has left organization respectively

---

#### **Distribution of Variables : Histogram**

```{r, echo=FALSE, warning=FALSE}
library(ggplot2)
attach(HR_Analytics)
ggplot(HR_Analytics,aes(x=satisfaction_level)) +
  facet_wrap(~ left)+
  geom_histogram(aes(y=..density..),col = "blue2",bins = 40) +
  stat_function(fun = dnorm,args = list(mean=mean(last_evaluation),sd=sd(last_evaluation)),colour = "red") +
  labs(title = "Distribution of Satisfaction level", x = "left", y = "Satisfaction level") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```

```{r, echo=FALSE, warnings=FALSE}
library(psych)
print(describeBy(HR_Analytics$satisfaction_level,group = HR_Analytics$left)[1:2])
```

* Distribution of Satisfaction levels across employees who retained & left in/from the organization appears normal

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics,aes(x=last_evaluation)) +
  facet_wrap(~ left)+
  geom_histogram(aes(y=..density..),col = "blue2", bins = 40) +
  stat_function(fun = dnorm,args = list(mean=mean(last_evaluation),sd=sd(last_evaluation)),colour = "red") +
  labs(title = "Distribution of Last Evaluation", x = "left", y = "Last Evaluation") +
  theme(plot.title = element_text(hjust = 0.50))
detach(HR_Analytics)
```

```{r, echo=FALSE, warnings=FALSE}
print(describeBy(HR_Analytics$last_evaluation,group = HR_Analytics$left)[1:2])
```

* Distribution of Last Evaluation across employees who retained & left in/from the organization appears normal

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics,aes(x=number_project)) +
  facet_wrap(~ left)+
  geom_histogram(aes(y=..density..),col = "blue2", bins = 20) +
  stat_function(fun = dnorm,args = list(mean=mean(number_project),sd=sd(number_project)),colour = "red") +
  labs(title = "Distribution of Number of Projects", x = "left", y = "Number of Project") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```

```{r, echo=FALSE, warnings=FALSE}
print(describeBy(HR_Analytics$number_project,group = HR_Analytics$left)[1:2])
```

* Distribution of Number of Projects across employees who retained & left in/from the organization appears normal

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics,aes(x=average_monthly_hours)) +
  facet_wrap(~ left)+
  geom_histogram(aes(y=..density..),col = "blue2", bins = 40) +
  stat_function(fun = dnorm,args = list(mean=mean(average_monthly_hours),sd=sd(average_monthly_hours)),
                colour = "red") +
  labs(title = "Distribution of Average Monthly Hours", x = "left", y = "Average Monthly Hours") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```

```{r, echo=FALSE, warnings=FALSE}
print(describeBy(HR_Analytics$average_monthly_hours,group = HR_Analytics$left)[1:2])
```

* Distribution of Average Monthly hours across employees who retained & left in/from the organization appears normal

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics,aes(x=time_spend_company)) +
  facet_wrap(~ left)+
  geom_histogram(aes(y=..density..),col = "blue2", bins = 40) +
  stat_function(fun = dnorm,args = list(mean=mean(time_spend_company),sd=sd(time_spend_company)),colour = "red") +
  labs(title = "Distribution of Time Spend in Company", x = "left", y = "Time Spend in Company") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```

```{r, echo=FALSE, warnings=FALSE}
print(describeBy(HR_Analytics$time_spend_company,group = HR_Analytics$left)[1:2])
```

* Distribution of Time Spend in Company of the employees who retained appears right skewed & who left appears normal

---

#### **Distribution of Variables : Boxplot**

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics,aes(left,satisfaction_level)) +
  geom_boxplot() +
  labs(title = "Distribution of Satisfaction level", x = "left", y = "Satisfaction level") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```

* Distribution of Satisfaction levels across employees who retained & left in/from the organization appears normal

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics,aes(left,last_evaluation)) +
  geom_boxplot() +
  labs(title = "Distribution of Last Evaluation", x = "left", y = "Last Evaluation") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```
* Distribution of Last Evaluation across employees who retained & left in/from the organization appears normal

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics,aes(left,number_project)) +
  geom_boxplot() +
  labs(title = "Distribution of Number of Projects", x = "left", y = "Number of Project") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```
* 50th & 75th percentile of employees who retained shares same number of projects with one outlier(6)
* Distribution of Number of Projects across employees who left from the organization appears normal

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics,aes(left,average_monthly_hours)) +
  geom_boxplot() +
  labs(title = "Distribution of Average Monthly Hours", x = "left", y = "Average Monthly Hours") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```
* Distribution of Average Monthly Hours across employees who retained & left in/from the organization appears normal

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics,aes(left,time_spend_company)) +
  geom_boxplot() +
  labs(title = "Distribution of Time Spend in Company", x = "left", y = "Time Spend in Company") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```
* Distribution of Average Monthly Hours across employees who retained appeared right skewed with outliers
* Distribution of Average Monthly Hours across employees who left appears normal

---

#### **Relationship between Variables : Scatterplot**

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics, aes(last_evaluation,satisfaction_level)) +
  facet_wrap(~left) +
  geom_point(aes(color=left)) +
  geom_abline(color = "red") +
  labs(title="Relationship between Satisfaction level & Last Evaluation", 
       x = "last evaluation", y = "satisfaction level") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```
```{r, echo=FALSE, warnings=FALSE}
library(data.table)
HR_Analytics <- data.table(HR_Analytics)
print(HR_Analytics[,.(cor(satisfaction_level, last_evaluation)), by = left])
HR_Analytics <- data.frame(HR_Analytics)
print(cor(HR_Analytics$satisfaction_level,HR_Analytics$last_evaluation))
```

* There is Weak relationship between Satisfaction level & Last evaluation across employees who retained & left in/from the Organization

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics, aes(scale(HR_Analytics$number_project)[,1],scale(HR_Analytics$satisfaction_level)[,1])) +
  facet_wrap(~left) +
  geom_point(aes(color=left)) +
  geom_abline(color = "red") +
  labs(title="Relationship between Satisfaction level & Number of Projects", 
       x = "Number of Projects", y = "satisfaction level") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```

```{r, echo=FALSE, warnings=FALSE}
HR_Analytics <- data.table(HR_Analytics)
print(HR_Analytics[,.(cor(satisfaction_level, number_project)), by = left])
HR_Analytics <- data.frame(HR_Analytics)
print(cor(HR_Analytics$satisfaction_level,HR_Analytics$number_project))
```

* There is Weak relationship between Satisfaction level & Number of Projects across employees who retained & left in/from the Organization

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics, aes(scale(average_monthly_hours)[,1],scale(satisfaction_level)[,1])) +
  facet_wrap(~left) +
  geom_point(aes(color=left)) +
  geom_abline(color = "red") +
  labs(title="Relationship between Satisfaction level & Average Monthly Hours", 
       x = "average monthly hours", y = "satisfaction level") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```

```{r, echo=FALSE, warnings=FALSE}
HR_Analytics <- data.table(HR_Analytics)
print(HR_Analytics[,.(cor(satisfaction_level, average_monthly_hours)), by = left])
HR_Analytics <- data.frame(HR_Analytics)
print(cor(HR_Analytics$satisfaction_level,HR_Analytics$average_monthly_hours))
```

* There is Weak relationship between Satisfaction level & Average Monthly Hours across employees who retained & left in/from the Organization

---

```{r, echo=FALSE, warnings=FALSE}
attach(HR_Analytics)
ggplot(HR_Analytics, aes(scale(time_spend_company)[,1],scale(satisfaction_level)[,1])) +
  facet_wrap(~left) +
  geom_point(aes(color=left)) +
  geom_abline(color = "red") +
  labs(title="Relationship between Satisfaction level & Time Spend in Company", 
       x = "time spend in company", y = "satisfaction level") +
  theme(plot.title = element_text(hjust = 0.5))
detach(HR_Analytics)
```

```{r, echo=FALSE, warnings=FALSE}
HR_Analytics <- data.table(HR_Analytics)
print(HR_Analytics[,.(cor(satisfaction_level, time_spend_company)), by = left])
HR_Analytics <- data.frame(HR_Analytics)
print(cor(HR_Analytics$satisfaction_level,HR_Analytics$time_spend_company))
```

* There is Weak relationship between Satisfaction level & Time Spend across employees who retained in the Organization
* There is fairly strong relationship between Satisfaction level & Time Spend across employees who left from Organization

---

#### **Relationship between Variables(Correlation Analysis) : Tetrachoric & Polychoric Correlation** [Reference](http://john-uebersax.com/stat/tetra.htm#tsoft)

```{r, echo=FALSE, warnings=FALSE}
library(polycor)
polychor(table(HR_Analytics$left,HR_Analytics$Work_accident), ML = T, std.err = T)
```
* There is fairly negative relationship between employees who left & work accident

---

```{r, echo=FALSE, warnings=FALSE}
polychor(table(HR_Analytics$left,HR_Analytics$promotion_last_5years), ML = T, std.err = T)
```
* There is fairly negative relationship between employees who left & given promotion in last 5-years

---

```{r, echo=FALSE, warnings=FALSE}
polychor(table(HR_Analytics$left,HR_Analytics$salary), ML = T, std.err = T)
```
* There is weak negative relationship between employees who left & salary break-up

---

#### **Correlation Matrix**

```{r, echo=FALSE, warnings=FALSE}
cor(HR_Analytics[sapply(HR_Analytics,is.numeric)])
```
* In general, there is no strong correlation exists between variables

---

#### **ANOVA**
##### Null Hypothesis: There is NO significant relationship between means(SL) of groups of left & other attributes
##### Alternate Hypothesis: There is significant relationship between means(SL) of groups of left & other attributes

```{r, echo=FALSE, warnings=FALSE}
anova(lm(satisfaction_level ~ Work_accident+promotion_last_5years+salary, data = HR_Analytics))
```

```{r, echo=FALSE, warnings=FALSE}
TukeyHSD(aov(satisfaction_level ~ Work_accident+promotion_last_5years+salary, data = HR_Analytics))
```

##### **Insights**

* There is a significant difference in the means of satisfaction level against work accident, promotion last 5-years & salary(low-high,medium-low)
* However, there is only 84% significant difference in the means of satisfaction level against medium-high salary bucket

---

#### **Principal Component Analysis**

```{r, echo=FALSE, warning=FALSE}
# PCA
PCA <- prcomp(HR_Analytics[sapply(HR_Analytics,is.numeric)][2:5],cor=TRUE, scale. = TRUE, retx = TRUE);PCA
summary(PCA)

#plot of variance of each PCA. It will be useful to decide how many principal components should be retained.
screeplot(PCA, type="lines",col=3) # 3-components can be retained

#biplot of first two principal components
biplot(PCA,cex=0.8)
abline(h = 0, v = 0, lty = 2, col = 8)
title("Bi-Plot")

# Identify which variable to drop
library(raster)
print(t(lapply(HR_Analytics[sapply(HR_Analytics,is.numeric)],sd)))
print(t(lapply(HR_Analytics[sapply(HR_Analytics,is.numeric)],cv)))

```

* Drop time_spend_company variable & continue transformation with first-3-components

* Note: As there are only few Variables, PCA is unable/not-suitable to explain more variances. Hence regression might not be prefered on PCA loadings

---

#### **Linear Regression**
##### Null Hypothesis: There is NO significant relationship between satisfaction level & other attributes
##### Alternate Hypothesis: There is significant relationship between satisfaction level & other attributes

##### **Simple/Stratified Random Sampling**

```{r, echo=FALSE, warnings=FALSE}
library(sampling)
HR_Analytics_Trans <- data.frame(PCA$x[,1:3])
names(HR_Analytics_Trans) <- names(HR_Analytics)[2:4]
HR_Analytics_Trans_Final <- cbind(satisfaction_level=HR_Analytics$satisfaction_level,HR_Analytics_Trans,
                                  HR_Analytics[c("Work_accident","promotion_last_5years","salary")])

HR_Analytics_Trans_Final_FT <- data.frame(table(HR_Analytics_Trans_Final$salary))
HR_Analytics_Trans_Final_FT$Per <- (HR_Analytics_Trans_Final_FT$Freq/sum(HR_Analytics_Trans_Final_FT$Freq))*100
names(HR_Analytics_Trans_Final_FT)[1] <- "salary"

# Consider 70% of the data as sample
HR_Analytics_Trans_Final_FT$Strata_Size <- ceiling((HR_Analytics_Trans_Final_FT$Freq*(ceiling((dim(HR_Analytics_Trans_Final)[1]/100)*70)/sum(HR_Analytics_Trans_Final_FT$Freq))))
HR_Analytics_Trans_Final_FT <- with(HR_Analytics_Trans_Final_FT,HR_Analytics_Trans_Final_FT[order(Strata_Size,decreasing = TRUE),])

# Stratification
HR_Analytics_Trans_Final_Strata <- strata(HR_Analytics_Trans_Final,c("salary"),
                                          size = HR_Analytics_Trans_Final_FT$Strata_Size, method = "srswor")
HR_Analytics_Trans_StRS <- getdata(HR_Analytics_Trans_Final,HR_Analytics_Trans_Final_Strata)
head(HR_Analytics_Trans_StRS)
```

##### **Linear Model** [Reference](http://stattrek.com/regression/linear-transformation.aspx?Tutorial=AP)

```{r, echo=FALSE, warnings=FALSE}
# Linear Model
summary(lm(satisfaction_level ~ last_evaluation+number_project+average_monthly_hours+Work_accident+
             promotion_last_5years+salary, data = HR_Analytics_Trans_StRS))
par(mfrow = c(2,2))
plot(lm(satisfaction_level ~ last_evaluation+number_project+average_monthly_hours+Work_accident+
             promotion_last_5years+salary, data = HR_Analytics_Trans_StRS))
```

---

```{r, echo=FALSE, warnings=FALSE}
# Transformation methods: Linear Model
summary(lm(log(satisfaction_level) ~ last_evaluation+number_project+average_monthly_hours+Work_accident+
             promotion_last_5years+salary, data = HR_Analytics_Trans_StRS))
par(mfrow = c(2,2))
plot(lm(exp(satisfaction_level) ~ last_evaluation+number_project+average_monthly_hours+Work_accident+
             promotion_last_5years+salary, data = HR_Analytics_Trans_StRS))
```

##### **Insights**

* Log-transformation is successful. However, overall varaitions explained by the model is just ~10%(~5%). Hence model will have poor prediction power
* Since p<0.05, we reject NH & conclude that there is a signficant relationship between satisfaction level & other attributes